Mermory-efficient Fujisaki-Okamoto transformation

### 2024-05-23 lattices/compact-fo-using-mac/manuscript/main.tex
Completed draft proof

    Under the random oracle model, game 3 and game 2 are indistinguishable from the adversary's perspective unless it makes a hash query on the value of $m^\ast$. Denote the probability that the adversary makes a hash query on $m^\ast$ by $P[\operatorname{QUERY}\ast]$, then:

    \begin{equation*}
        \epsilon_2 - \epsilon_3 \leq P[\operatorname{QUERY}\ast]
    \end{equation*}

    we can bound $P[\operatorname{QUERY}\ast]$ by constructing an OW-CPA adversary that simulates game 3 for a OW-PCVA adversary as a sub-routine. After the OW-PCVA adversary halts, the OW-CPA adversary checks through the tape of the hash oracle. Since the encryption routine is deterministic, if $m^\ast$ exists in the hash oracle tape, then it can be identified for sure. In other words, if $m^\ast$ has been queried by the OW-PCVA sub-routine, then the OW-CPA adversary can be guaranteed to win its game

    \begin{equation*}
        P[\operatorname{QUERY}\ast] = \epsilon_\text{OW-CPA}
    \end{equation*}

@@

\begin{theorem}
    For every OW-CPA adversary against $(K, E^{\strike{\$}}), D$ that makes $q_H$ hash queries to $H$ and has advantage $\epsilon_{\strike{\$}}$, there exists an OW-CPA adversary against $(K, E, D)$ with advantage $\epsilon_\text{OW-CPA}$ such that:

    \begin{equation*}
        \epsilon_{\strike{\$}} \leq (1 + q_H) \cdot \epsilon_\text{OW-CPA}
    \end{equation*}
\end{theorem}

\begin{proof}
    % Game 1: replace pseudorandom coin with random coin and use QUERY* argument, this time we will lose q_H * epsilon amount of security, probably

    % Game 1 can be simulated by OW-CPA adversary
\end{proof}

>>>>

    Under the random oracle model, game 3 and game 2 are indistinguishable from the adversary's perspective unless it makes a hash query on the value of $m^\ast$. Denote the probability that the adversary makes a hash query on $m^\ast$ by $P[\operatorname{QUERY}^\ast]$, then:

    \begin{equation*}
        \epsilon_2 - \epsilon_3 \leq P[\operatorname{QUERY}^\ast]
    \end{equation*}

    we can bound $P[\operatorname{QUERY}^\ast]$ by constructing an OW-CPA adversary that simulates game 3 for a OW-PCVA adversary as a sub-routine. After the OW-PCVA adversary halts, the OW-CPA adversary checks through the tape of the hash oracle. Since the encryption routine is deterministic, if $m^\ast$ exists in the hash oracle tape, then it can be identified for sure. In other words, if $m^\ast$ has been queried by the OW-PCVA sub-routine, then the OW-CPA adversary can be guaranteed to win its game

    \begin{equation*}
        P[\operatorname{QUERY}^\ast] = \epsilon_\text{OW-CPA}
    \end{equation*}

@@

\begin{theorem}
    For every OW-CPA adversary against the deterministic PKE $(K, E^{\strike{\$}}), D$ that makes $q_H$ hash queries to $H$ and has advantage $\epsilon_{\strike{\$}}$, there exists an OW-CPA adversary against the probabilistic PKE $(K, E, D)$ with advantage $\epsilon_\$$ such that:

    \begin{equation*}
        \epsilon_{\strike{\$}} \leq (1 + q_H) \cdot \epsilon_\$
    \end{equation*}
\end{theorem}

\begin{proof}
    We prove using a sequence of game. Denote the OW-CPA adversary against the deterministic PKE by $\mathcal{A}_{\strike{\$}}$ and its advantage by $\epsilon_{\strike{\$}}$; denote the OW-CPA adversary against the probabilistic PKE by $\mathcal{A}_\$$ and its advantage by $\epsilon_\$$

    Game 0 is the standard OW-CPA game: $\epsilon_0 = \epsilon_{\strike{\$}}$

    In game 1, the challenge encryption routine is modified. Instead of using a pseudorandom coin, a truly random coin is used.

    \begin{multicols}{2}
        \begin{algorithm}[H]
            \caption{Game 0}
            \SetAlgoLined
            $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
            $m^\ast \leftsample \mathcal{M}$\;
            $r^\ast \leftarrow H(m^\ast)$\;
            $c^\ast \leftarrow E(\pk, m^\ast; r^\ast)$\;
            $\hat{m} \leftsample \mathcal{A}_\text{OW-CPA}(1^\lambda, \pk, c^\ast)$\;
            \Return{
                $\llbrack \hat{m} = m^\ast\rrbrack$
            }
        \end{algorithm}

        \begin{algorithm}[H]
            \caption{Game 1}
            \SetAlgoLined
            $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
            $m^\ast \leftsample \mathcal{M}$\;
            $r^\ast \leftsample \coin$\;
            $c^\ast \leftarrow E(\pk, m^\ast; r^\ast)$\;
            $\hat{m} \leftsample \mathcal{A}_\text{OW-CPA}(1^\lambda, \pk, c^\ast)$\;
            \Return{
                $\llbrack \hat{m} = m^\ast\rrbrack$
            }
        \end{algorithm}
    \end{multicols}

    Under the random oracle assumption, game 0 and game 1 are indistinguishable from the perspective of $\mathcal{A}_{\strike{\$}}$ except for when $\mathcal{A}_{\strike{\$}}$ makes a hash query $\tilde{m} = m^\ast$. Denote the event that $\mathcal{A}_{\strike{\$}}$ makes such a query by $\operatorname{QUERY}^\ast$, then by the difference lemma:

    \begin{equation*}
        \epsilon_0 - \epsilon_1 \leq P[\operatorname{QUERY}^\ast]
    \end{equation*}

    We can bound $P[\operatorname{QUERY}^\ast]$ by constructing a OW-CPA adversary against the probabilistic scheme that uses $\mathcal{A}_{\strike{\$}}$ as a sub-routine and simulates game 1 for $\mathcal{A}_{\strike{\$}}$. When $\mathcal{A}_\$$ receives the probabilistic challenge ciphertext $c^\ast$, it directly passes it to the sub-routine $\mathcal{A}_{\strike{\$}}$. After the sub-routine halts, $\mathcal{A}_\$$ samples from the hash oracle tape a random message as its output. If the sub-routine $\mathcal{A}_{\strike{\$}}$ indeed makes the correct query $m^\ast$, then the probability that a randomly chosen query value being the correct value is $\frac{1}{q_H}$:

    \begin{equation*}
        \epsilon_\$ = P[\operatorname{QUERY}^\ast] \cdot \frac{1}{q_H}
    \end{equation*}

    Game 1 can be simulated by an OW-CPA adversary against the probabilistic PKE:

    \begin{equation*}
        \epsilon_1 = \epsilon_\$
    \end{equation*}

    Putting the three equations above together gives us the desired inequality.
\end{proof}
